Supplemental Online Materials:\\\textit{Double Standards in Judging Collective Action}

### 2022-05-24 Documents/supplemental-materials.tex
work in progress

---
title: Supplemental Online Materials
subtitle: Double Standards in Judging Collective Action
bibliography: references.bib
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: som-template.tex
csl: apa.csl
---

```{r setup, include = FALSE}

  # Load packages
  library(tidyverse); library(tidybayes); library(numform)

  # knitr options
  knitr::opts_chunk$set(echo = FALSE)

```

# Scale Development

## Study 1

@@

> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong. Each person lives in a place, has a job (or not) at an organisation, is a fan of a specific sports club,  has a religion, or belongs to any number of other such groups. Individuals often act in ways to promote the interests of the social groups to which they belong.

> For example, the workers in a factory might want to get paid more. They might go on a strike to reach that goal. Students at a university might want to prevent an increase in tuition fees. They might hand out flyers or occupy a building to reach that goal. Other groups might want to protest against police violence. They might block traffic on a road to make others aware of this issue.

@@

> Please think of other actions that members of a social group (or social groups as a whole) might take to promote their group's interests or goals. Please name at least 5 actions that fit the description above. You can name up to 10 actions if you want to. Please try to think of actions that vary in how acceptable or unacceptable they are (in your opinion).

\noindent After naming at least five actions, participants sorted their responses into actions that, across a range of situations, were "always acceptable", "sometimes acceptable", and "never acceptable":

> On the left, you see the actions you have named on the previous page. Think about each action. Can you think of situations in which this action is an acceptable means to advance a group's goals or interests? Can you think of situations in which this action is not an acceptable means to advance a group's interests?

\noindent Participants named between 5 and 10 actions ($\textit{Mdn} = 7$). Participants named more actions that they considered always acceptable ($\textit{Mdn} = 3$) than actions they considered sometimes ($\textit{Mdn} = 2$) or never ($\textit{Mdn} = 1$) acceptable. We recoded participants' responses into a smaller set of unique collective actions. We then supplemented the resulting list of actions with collective actions from the psychological and political science literature [e.g., @sharp_nonviolent_1973]. This process resulted in 72 actions that we expected to vary in how acceptable most people would find them to be.

## Study 2

In Study 2, we measured how acceptable participants judged the actions from Study 1 to be and applied item response theory to develop an instrument to capture double standards in judging collective action. We recruited 158 participants ($\textit{Mdn} = 30$ years, age range: 18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the Prolific subject pool, all of whom were citizens of the UK or the US. To increase the socioeconomic diversity of our sample, we recruited 80 non-students without a university degree, 37 non-students with a university degree, and 41 current university students. We excluded 15 participants who failed an attention check, leaving a final sample of 143 participants for our analyses.

@@

> Society is not only made up of individuals, but consists of 'social groups' to which these individuals belong. Each person lives in a place, has a job (or not) at a firm or organisation, has a religion (or not), is part of a political party, or belongs to any number of other groups like these.

> Individuals often act in ways to promote their group's interest. For example, individuals might want people like them to get paid more. Members of a group might take various kinds of actions (for example, go on a strike).

\noindent Participants were then asked to "suppose that one or more members of a group took the following action to advance a cause in their group's interest" and were presented with one of the actions from Study 1. Participants answered several questions about the protest action. First, they rated how disruptive, violent, and extreme they considered this action to be (1 = *not at all*, 4 = *very*). Second, they were asked to think of a range of different causes and circumstances, and to rate how often this action would be an acceptable means for a group to advance one of these causes (1 = *never*, 2 = *rarely*, 3 = *sometimes*, 4 = *often*, 5 = *always*). Finally, they rated how positive or negative they felt, in general, about this action (1 = *very positive*, 5 = *very negative*). Each participant answered these questions for 20 of the 72 actions from Study 1 so that each action was rated by 29--53 participants.

We estimated a graded response model [@burkner_bayesian_2021; @samejima_graded_1997], an item response theory model for ordinal response variables, for participants' ratings of how often an action would be an acceptable form of collective action. For each participant $j$, the model estimated their unique propensity ($\theta_j$) to consider collective actions acceptable in more situations. For each item $i$, the model estimated four acceptability thresholds ($\beta_{ik}$), separating the five response options, and one discrimination parameter ($\alpha_i$) indicating how well the item differentiates between participants with different propensities to consider collective actions acceptable. We estimated the model in *CmdStanR* using similar prior distribution as in Experiment. We used an induced Dirichlet prior [@betancourt_ordinal_2019] for the item-specific difficulty thresholds.

@@

# Experiment 1

## Method

### Procedure

@@

First, we recruited a larger pool of 875 participants. Of these, 475 participants did not have a university degree and placed themselves on the bottom three ranks of the subjective socio-economic status ladder (*lower-status group*) and 400 participants had at least an undergraduate degree and placed themselves on the top four ranks of the subjective socio-economic status ladder (*higher-status group*). Participants completed a screening survey in which they, among other questions, answered whether they considered either their current job or the jobs they had had in the past or would have in the future to be a working-class job or a middle-class/professional job (1 = *working-class job*, 2 = *middle-class/professional job*, 3 = *neither*). Before answering this question, participants read the following explanation:

> The workforce is often divided into two kinds of jobs. What we call working-class jobs are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers. These jobs do not usually require a university degree. What we call middle-class or professional jobs are administrative, managerial, or other jobs that usually require a university degree.

\noindent As preregistered, we excluded participants from the lower-status group who had responded "middle-class/professional job" or "neither" to this question; we excluded participants from the higher-status group who had responded "working-class job" or "neither" to this question. This left 687 participants for the next step of the selection procedure.

Second, we set out to recruit 500 participants from the remaining 687 participants, 250 from the lower-status group and 250 from the higher-status group. After giving their informed consent, participants were randomly assigned to read a vignette about a government bill affecting either people in working-class jobs (*lower-status actors*) or people in professional jobs (*higher-status actors*). Participants in both conditions were instructed to carefully read the vignette and to try to imagine what it would be like if this situation was real. Participants in the lower-status actors condition then read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in working-class jobs. Working-class jobs, in this case, are jobs done by skilled, semi-skilled, unskilled manual workers or by casual workers. These are jobs that do not usually require a university degree. Other jobs are unlikely to be affected.
Participants in the higher-status actors condition instead read the following introduction:

> The government, though not necessarily the current government, is going to introduce a bill that will mostly affect people in professional jobs. Professional jobs are administrative, managerial, or other jobs that usually require a university degree. Other jobs are unlikely to be affected.
Participants in the two conditions then read this almost identically worded paragraph:

>This government measure would make it easier for companies to hire workers during economic growth and to lay off workers during an economic crisis. As a consequence, companies would be able to fire employees with little notice and without giving a reason. Trade unions are opposed to the measure. They argue that the bill would compromise job security, and prevent employees from challenging harassment or other abuse without the fear of being fired. People in [working-class/professional] jobs are particularly at risk, and there is a rise in tension and outrage among them.

@@

# References

@@

<div id="refs"></div>

@@

```{=latex}
\begin{xltabular}{\linewidth}{rXrrrrrrr}

@@

\end{xltabular}
```

@@

```{=latex}
\setcounter{table}{1}

@@

```{=latex}
\setcounter{table}{2}

\begin{table}[t!]
\caption{Correlations between ratings in Study 2}
\centering
\begin{tabular}{rlrrrrr}
\toprule
\# & Rating & Acceptable & Disruptive & Violent & Extreme & Negative\\
\midrule1 & Acceptable &  & -.69 & -.64 & -.77 & -.89\\
2 & Disruptive & -.69 &  & .66 & .78 & .70\\
3 & Violent & -.64 & .66 &  & .79 & .64\\
4 & Extreme & -.77 & .78 & .79 &  & .77\\
5 & Negative & -.89 & .70 & .64 & .77 & \\
\bottomrule
\addlinespace
\end{tabular}
\caption*{\textit{Note.} Acceptable (1 = \textit{never}, 5 = \textit{always}); Disruptive, Violent, Extreme (1 = \textit{not at all}, 4 = \textit{very}); Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\end{table}
```

>>>>

\documentclass[12pt, letterpaper]{article}

% Font
\usepackage[utf8]{inputenc}
\usepackage{MinionPro} 
\input{glyphtounicode}
\pdfgentounicode=1 
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setcounter{secnumdepth}{0}
\usepackage{titlesec}  
\titleformat*{\section}{\centering\normalfont\Large\bfseries}

% Links
\usepackage[colorlinks = true, linkcolor = black, urlcolor = black, citecolor = black]{hyperref}

% Figures
\usepackage{graphicx}
\usepackage[labelfont = bf, font = small, labelsep = newline, singlelinecheck = false]{caption}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

% Tables
\usepackage{booktabs}
\usepackage{xltabular}

% References
\newenvironment{CSLReferences}[2]{}{}

% Frontmatter
\title{ Supplemental Online Materials:\\\textit{Double Standards in
Judging Collective Action} }
\author{  }

\begin{document}

\maketitle

\hypertarget{scale-development}{%
\section{Scale Development}\label{scale-development}}

\hypertarget{study-1}{%
\subsection{Study 1}\label{study-1}}

@@

\begin{quote}
Society is not only made up of individuals, but consists of `social
groups' to which these individuals belong. Each person lives in a place,
has a job (or not) at an organisation, is a fan of a specific sports
club, has a religion, or belongs to any number of other such groups.
Individuals often act in ways to promote the interests of the social
groups to which they belong.
\end{quote}

\begin{quote}
For example, the workers in a factory might want to get paid more. They
might go on a strike to reach that goal. Students at a university might
want to prevent an increase in tuition fees. They might hand out flyers
or occupy a building to reach that goal. Other groups might want to
protest against police violence. They might block traffic on a road to
make others aware of this issue.
\end{quote}

@@

\begin{quote}
Please think of other actions that members of a social group (or social
groups as a whole) might take to promote their group's interests or
goals. Please name at least 5 actions that fit the description above.
You can name up to 10 actions if you want to. Please try to think of
actions that vary in how acceptable or unacceptable they are (in your
opinion).
\end{quote}

\noindent After naming at least five actions, participants sorted their
responses into actions that, across a range of situations, were ``always
acceptable'', ``sometimes acceptable'', and ``never acceptable'':

\begin{quote}
On the left, you see the actions you have named on the previous page.
Think about each action. Can you think of situations in which this
action is an acceptable means to advance a group's goals or interests?
Can you think of situations in which this action is not an acceptable
means to advance a group's interests?
\end{quote}

\noindent Participants named between 5 and 10 actions
(\(\textit{Mdn} = 7\)). Participants named more actions that they
considered always acceptable (\(\textit{Mdn} = 3\)) than actions they
considered sometimes (\(\textit{Mdn} = 2\)) or never
(\(\textit{Mdn} = 1\)) acceptable. We recoded participants' responses
into a smaller set of unique collective actions. We then supplemented
the resulting list of actions with collective actions from the
psychological and political science literature (e.g., Sharp, 1973). This
process resulted in 72 actions that we expected to vary in how
acceptable most people would find them to be.

\hypertarget{study-2}{%
\subsection{Study 2}\label{study-2}}

In Study 2, we measured how acceptable participants judged the actions
from Study 1 to be and applied item response theory to develop an
instrument to capture double standards in judging collective action. We
recruited 158 participants (\(\textit{Mdn} = 30\) years, age range:
18--68 years; 103 women, 52 men, 2 other, 1 prefer not to say) from the
Prolific subject pool, all of whom were citizens of the UK or the US. To
increase the socioeconomic diversity of our sample, we recruited 80
non-students without a university degree, 37 non-students with a
university degree, and 41 current university students. We excluded 15
participants who failed an attention check, leaving a final sample of
143 participants for our analyses.

@@

\begin{quote}
Society is not only made up of individuals, but consists of `social
groups' to which these individuals belong. Each person lives in a place,
has a job (or not) at a firm or organisation, has a religion (or not),
is part of a political party, or belongs to any number of other groups
like these.
\end{quote}

\begin{quote}
Individuals often act in ways to promote their group's interest. For
example, individuals might want people like them to get paid more.
Members of a group might take various kinds of actions (for example, go
on a strike).
\end{quote}

\noindent Participants were then asked to ``suppose that one or more
members of a group took the following action to advance a cause in their
group's interest'' and were presented with one of the actions from Study
1. Participants answered several questions about the protest action.
First, they rated how disruptive, violent, and extreme they considered
this action to be (1 = \emph{not at all}, 4 = \emph{very}). Second, they
were asked to think of a range of different causes and circumstances,
and to rate how often this action would be an acceptable means for a
group to advance one of these causes (1 = \emph{never}, 2 =
\emph{rarely}, 3 = \emph{sometimes}, 4 = \emph{often}, 5 =
\emph{always}). Finally, they rated how positive or negative they felt,
in general, about this action (1 = \emph{very positive}, 5 = \emph{very
negative}). Each participant answered these questions for 20 of the 72
actions from Study 1 so that each action was rated by 29--53
participants.

We estimated a graded response model (Bürkner, 2021; Samejima, 1997), an
item response theory model for ordinal response variables, for
participants' ratings of how often an action would be an acceptable form
of collective action. For each participant \(j\), the model estimated
their unique propensity (\(\theta_j\)) to consider collective actions
acceptable in more situations. For each item \(i\), the model estimated
four acceptability thresholds (\(\beta_{ik}\)), separating the five
response options, and one discrimination parameter (\(\alpha_i\))
indicating how well the item differentiates between participants with
different propensities to consider collective actions acceptable. We
estimated the model in \emph{CmdStanR} using similar prior distribution
as in Experiment. We used an induced Dirichlet prior (Betancourt, 2019)
for the item-specific difficulty thresholds.

@@

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

\hypertarget{method}{%
\subsection{Method}\label{method}}

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

@@

First, we recruited a larger pool of 875 participants. Of these, 475
participants did not have a university degree and placed themselves on
the bottom three ranks of the subjective socio-economic status ladder
(lower-status group) and 400 participants had at least an undergraduate
degree and placed themselves on the top four ranks of the subjective
socio-economic status ladder (higher-status group). Participants
completed a screening survey in which they, among other questions,
answered whether they considered either their current job or the jobs
they had had in the past or would have in the future to be a
working-class job or a middle-class/professional job (1 = working-class
job, 2 = middle-class/professional job, 3 = neither). Before answering
this question, participants read the following explanation:

\begin{quote}
The workforce is often divided into two kinds of jobs. What we call
working-class jobs are jobs done by skilled, semi-skilled, unskilled
manual workers or by casual workers. These jobs do not usually require a
university degree. What we call middle-class or professional jobs are
administrative, managerial, or other jobs that usually require a
university degree.
\end{quote}

As preregistered, we excluded participants from the lower-status group
who had responded ``middle-class/professional job'' or ``neither'' to
this question; we excluded participants from the higher-status group who
had responded ``working-class job'' or ``neither'' to this question.
This left 687 participants for the next step of the selection procedure.

Second, we set out to recruit 500 participants from the remaining 687
participants, 250 from the lower-status group and 250 from the
higher-status group. After giving their informed consent, participants
were randomly assigned to read a vignette about a government bill
affecting either people in working-class jobs (lower-status actors) or
people in professional jobs (higher-status actors). Participants in both
conditions were instructed to carefully read the vignette and to try to
imagine what it would be like if this situation was real. Participants
in the lower-status actors condition then read the following
introduction:

\begin{quote}
The government, though not necessarily the current government, is going
to introduce a bill that will mostly affect people in working-class
jobs. Working-class jobs, in this case, are jobs done by skilled,
semi-skilled, unskilled manual workers or by casual workers. These are
jobs that do not usually require a university degree. Other jobs are
unlikely to be affected. Participants in the higher-status actors
condition instead read the following introduction:
\end{quote}

\begin{quote}
The government, though not necessarily the current government, is going
to introduce a bill that will mostly affect people in professional jobs.
Professional jobs are administrative, managerial, or other jobs that
usually require a university degree. Other jobs are unlikely to be
affected. Participants in the two conditions then read this almost
identically worded paragraph:
\end{quote}

\begin{quote}
This government measure would make it easier for companies to hire
workers during economic growth and to lay off workers during an economic
crisis. As a consequence, companies would be able to fire employees with
little notice and without giving a reason. Trade unions are opposed to
the measure. They argue that the bill would compromise job security, and
prevent employees from challenging harassment or other abuse without the
fear of being fired. People in {[}working-class/professional{]} jobs are
particularly at risk, and there is a rise in tension and outrage among
them.
\end{quote}

@@

\hypertarget{references}{%
\section{References}\label{references}}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-betancourt_ordinal_2019}{}}%
Betancourt, M. (2019). \emph{Ordinal regression}.
\url{https://betanalpha.github.io/assets/case_studies/ordinal_regression.html}

\leavevmode\vadjust pre{\hypertarget{ref-burkner_bayesian_2021}{}}%
Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with brms
and {Stan}. \emph{Journal of Statistical Software}, \emph{100}(5).
\url{https://doi.org/10.18637/jss.v100.i05}

\leavevmode\vadjust pre{\hypertarget{ref-samejima_graded_1997}{}}%
Samejima, F. (1997). Graded response model. In W. J. van der Linden \&
R. K. Hambleton (Eds.), \emph{Handbook of modern item response theory}
(pp. 85--100). Springer.

\leavevmode\vadjust pre{\hypertarget{ref-sharp_nonviolent_1973}{}}%
Sharp, G. (1973). \emph{The politics of nonviolent action}. Porter
Sargent.

\end{CSLReferences}

@@

\begin{xltabular}{\linewidth}{rXrrrrrrr}

@@

\end{xltabular}

@@

\setcounter{table}{1}

@@

\setcounter{table}{2}

\begin{table}[t!]
\caption{Correlations between ratings in Study 2}
\centering
\begin{tabular}{rlrrrrr}
\toprule
\# & Rating & Acceptable & Disruptive & Violent & Extreme & Negative\\
\midrule1 & Acceptable &  & -.69 & -.64 & -.77 & -.89\\
2 & Disruptive & -.69 &  & .66 & .78 & .70\\
3 & Violent & -.64 & .66 &  & .79 & .64\\
4 & Extreme & -.77 & .78 & .79 &  & .77\\
5 & Negative & -.89 & .70 & .64 & .77 & \\
\bottomrule
\addlinespace
\end{tabular}
\caption*{\textit{Note.} Acceptable (1 = \textit{never}, 5 = \textit{always}); Disruptive, Violent, Extreme (1 = \textit{not at all}, 4 = \textit{very}); Negative (1 = \textit{very positive}, 5 = \textit{very negative})}
\end{table}

\end{document}
